### 2019-08-21 Manuscript/Manuscript.tex
Updates from Overleaf

The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes. The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a, Hartmann2014a}. There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a}; Fig. \ref{Fig:indo_growth}). This collision between the Australian-Indian and Eurasian plates and intervening arc terranes has lead to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics.

Given that the region is a modern-day hotspot of chemical weathering and \COtwo consumption, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be on Earth if the region was not there at all? Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate state. There has been significant planetary cooling since the Mid-Miocene climatic optimum leading to the onset of northern hemisphere glaciation, in addition to Antarctic glaciation, over the past few million years \citep{Shackleton1984a}. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750 ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280 ppm \citep{DeConto2008a} providing an explanation for the delay between southern hemisphere and bipolar glaciation. We seek to evaluate the hypothesis that the growth of the broader Indonesian archipelago has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.

Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}. However, assessing the effect of the growth of the broader Indonesia archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes. As \COtwo sinks are removed and \COtwo rises, there will be increases in temperature and invigorated hydrological cycling that causes other weathering sinks to increase. The steady-state \COtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources. The magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.

@@

Topography, climate and lithology all play important roles in modulating chemical weathering. In low-relief regions, low physical erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \citep{Gabet2009a, West2012a, Maher2014a}. In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \citep{Lasaga1994a, West2012a}. Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance the total chemical weathering rate \citep{West2012a, Maher2014a}. In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}. These factors taken together are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability and initiating glacial climate over the past 520 million years \citep{Jagoutz2016a,Swanson-Hysell2017a,Macdonald2019a} and perhaps in the Neoproterozoic as well \citep{Park2019a}.

@@

The areas of Indonesia and New Guinea has substantially increased over this time period with an acceleration over the past 5 million years.

We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics. We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.

\subsection*{GEOCLIM}

@@

Geological maps and paleo-shoreline data were used to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea. Following Molnar and Cronin (10), we analyzed the area changes of islands that are greater than ~200 km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda shelf. Larger islands were further divided into regions based upon their position on microplates as defined by Matthews, et al. (78). Lithological maps were modified from Hartmann and Moosdorf (79) and then supplemented with regional geological maps. From these maps, the percentage of ultramafic and volcanic rocks relative to other rock types was calculated in QGIS for each island and microplate region. 

@@

We used geological maps and paleo-shoreline data to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea. Using Gplates, paleoshoreline shapefiles were incorporated in the paleogeographic model of Matthews, et al. (78). Although there are some differences between this reconstruction and the recent reconstruction of Hall (33), these do not entail significant changes in paleolatitude. 

>>>>

The broader Indonesia archipelago has an out-sized influence on modern chemical weathering fluxes. The confluence of steep topography, a warm and wet tropical climate, and the presence of mafic lithologies results in high fluxes of Ca and Mg cations and associated \COtwo consumption \citep{Hartmann2009a, Hartmann2014a}. There has been significant growth of land area within the region since the Miocene associated with ongoing arc-continent collision between Australia and the Sunda-Banda arc system (\citealp{Molnar2015a}; Fig. \ref{Fig:indo_growth}). This collision between the Australian-Indian and Eurasian plates and intervening arc terranes has led to the emergence of Indonesia and New Guinea and the exhumation of large supra-subduction ophiolites and associated forearcs in the warm, wet tropics.

Given that the region is a modern-day hot spot of chemical weathering and \COtwo consumption \citep{Hartmann2009a}, what effect has growth of the broader Indonesian archipelago had on Earth's climate state and what would \pCOtwo be on Earth if the region was not there at all? Based on paleogeographic reconstructions of arc-continent collisions through time, \citet{Macdonald2019a} proposed that arc-continent collisions in the tropics are responsible for increasing planetary weatherability, lowering \pCOtwo and causing the onset of glacial climate states. There has been significant planetary cooling since the Mid-Miocene Climatic Optimum leading to the onset of northern hemisphere glaciation, in addition to Antarctic glaciation, over the past few million years \citep{Shackleton1984a}. The \pCOtwo threshold for Antarctic glaciation is estimated to be $\sim$750~ppm with that for northern hemisphere glaciation being significantly lower at $\sim$280~ppm \citep{DeConto2008a} providing an explanation for the delay between southern hemisphere and bipolar glaciation. We seek to evaluate the hypothesis that the growth of the broader Indonesian archipelago has played a significant role in lowering \pCOtwo thereby driving this cooling and the development of northern hemisphere ice sheets.

Over geologic time-scales, \COtwo enters Earth's ocean/atmosphere system primarily via volcanism and metamorphic degassing of the solid Earth, and leaves it primarily through the chemical weathering of silicate rocks, which delivers alkalinity and cations to the ocean that ultimately result in the consumption of carbon via the precipitation of carbonate rocks \citep{Kump2000a}. However, assessing the effect of the growth of the broader Indonesia archipelago on Earth's \pCOtwo levels can not be determined simply through assessing the present-day chemical weathering fluxes. As \COtwo sinks are removed and \COtwo rises, there will be increases in temperature and invigorated hydrological cycling that causes other weathering sinks to increase. The steady-state \pCOtwo of Earth is set by the \pCOtwo level at which the sinks of \COtwo are equal to the sources. The magnitude of the silicate weathering \COtwo sink can be equal to the magnitude of the source at lower \pCOtwo on a planet with relatively high weatherability than on a planet where it is lower \citep{Kump1997a}.

@@

Topography, climate, and lithology all play important roles in modulating chemical weathering. In low-relief regions, low physical erosion rates lead to a supply-limited weathering regime, in contrast to regions of high denudation where there are higher weathering fluxes \citep{Gabet2009a, West2012a, Maher2014a}. In warm and wet regions, mineral dissolution kinetics are faster leading to enhanced chemical weathering \citep{Lasaga1994a, West2012a}. Regions with high physical weathering rates that are not supply-limited are more sensitive to changes in climate that can enhance the total chemical weathering rate \citep{West2012a, Maher2014a}. In terms of lithology, mafic rocks have the potential to sequester more carbon through silicate weathering because of their relatively high concentrations of Ca and Mg as well as high reaction rates \citep{Dessert2003a}. These factors taken together are what have lead to the proposal that arc-continent collisions within the tropical rain belt have been important in enhancing global weatherability and initiating glacial climate over the past 520~m.y. \citep{Jagoutz2016a, Swanson-Hysell2017a, Macdonald2019a} and perhaps in the Neoproterozoic as well \citep{Park2019a}.

@@

%The areas of Indonesia and New Guinea has substantially increased over this time period with an acceleration over the past 5 million years.

%We then incorporate these rock units into a paleogeographic model to determine the position of these units relative to the warm wet tropics. We find that the subaerially exposed area in the tropics of Indonesia and New Guinea has steadily increased through the Neogene with an acceleration over the past 5 Myrs.

% \subsection*{GEOCLIM}

@@

Geological maps and paleoshoreline data were used to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea. Following Molnar and Cronin (10), we analyzed the area changes of islands that are greater than ~200 km, and we also included changes in areas of submerged continental shelves that were previously exposed, like the Sunda shelf. Larger islands were further divided into regions based upon their position on microplates as defined by Matthews, et al. (78). Lithological maps were modified from Hartmann and Moosdorf (79) and then supplemented with regional geological maps. From these maps, the percentage of ultramafic and volcanic rocks relative to other rock types was calculated in QGIS for each island and microplate region.

@@

We used geological maps and paleoshoreline data to calculate the changes in different types of subaerially exposed rocks in Indonesia and New Guinea. Using Gplates, paleoshoreline shapefiles were incorporated in the paleogeographic model of Matthews, et al. (78). Although there are some differences between this reconstruction and the recent reconstruction of Hall (33), these do not entail significant changes in paleolatitude.
